Final project

### 2025-04-06 code/eda-code/eda.qmd
Made a few changes to the readme files

# Library/install libraries
```{r}
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(corrplot)
```

# EXPLORATORY DATA ANALYSIS (EDA)
```{r}
# Import the dataset
cardio_df <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

@@

# Calculate summary statistics for selected variables
```{r}
# Calculate summary statistics for selected variables
summary_stats <- cardio_df %>% 
  summarise(
    # Age statistics
    Age_min = min(age, na.rm = TRUE),
    Age_max = max(age, na.rm = TRUE),
    Age_mean = mean(age, na.rm = TRUE),
    Age_sd = sd(age, na.rm = TRUE),

@@

summary_df = skimr::skim(summary_stats)
print(summary_stats)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Age distribution
```{r}
p1 <- ggplot(cardio_df, aes(x = age)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "steelblue", alpha = 0.5) +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(title = "Age Distribution", x = "Age", y = "Density") +
  scale_x_continuous(breaks = seq(0, max(cardio_df$age, na.rm = TRUE) + 10, by = 10)) +
  theme_bw()

@@

# Serum Cholesterol distribution
```{r}
p2 <- ggplot(subset(cardio_df, serumcholestrol > 0), aes(x = serumcholestrol)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "forestgreen", alpha = 0.5) +
  geom_density(color = "darkgreen", linewidth = 1) +
  labs(title = "Serum Cholesterol Distribution", 
       x = "Serum Cholesterol (mg/dL)", y = "Density") +
  theme_minimal()

@@

# Resting Blood Pressure distribution
```{r}
p3 <- ggplot(subset(cardio_df, restingBP > 0), aes(x = restingBP)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "maroon", alpha = 0.5) +
  geom_density(color = "darkred", linewidth = 1) +
  labs(title = "Resting Blood Pressure Distribution", 
       x = "Resting Blood Pressure (mmHG)", y = "Density") +
  theme_bw()

@@

# Boxplot for resting blood pressure by heart disease status
```{r}
p4 <- ggplot(cardio_df, aes(x = target, y = restingBP, fill = target)) +
  geom_boxplot() +
  labs(title = "Resting Blood Pressure by Heart Disease Status",
       x = "Heart Disease Status",
       y = "Resting Blood Pressure (mmHg)") +
  theme_minimal()

@@

# Violin plot for oldpeak by heart disease status
```{r}
p5 <- ggplot(cardio_df, aes(x = target, y = oldpeak, fill = target)) +
  geom_violin(trim = FALSE) +
  labs(title = "Oldpeak Distribution by Heart Disease Status",
       x = "Heart Disease Status",
       y = "Oldpeak") +
  theme_minimal()

@@

# Create the correlation matrix for selected numeric variables
```{r}
numeric_vars <- cardio_df %>% 
  select(age, restingBP, serumcholestrol, maxheartrate, oldpeak)
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")

>>>>

# Library/install libraries

```{r}
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(corrplot)
```

# EXPLORATORY DATA ANALYSIS (EDA)

```{r}
# Import the dataset
cardio_df <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

@@

# Calculate summary statistics for selected variables

```{r}
# Calculate summary statistics for selected variables
summary_stats <- cardio_df %>% 
  summarise(
    # Age statistics
    Age_min = min(age, na.rm = TRUE),
    Age_max = max(age, na.rm = TRUE),
    Age_mean = mean(age, na.rm = TRUE),
    Age_sd = sd(age, na.rm = TRUE),

@@

summary_df = skimr::skim(summary_stats)
print(summary_stats)
# save to file
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary_df, file = summarytable_file)
```

# Age distribution

```{r}
p1 <- ggplot(cardio_df, aes(x = age)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "steelblue", alpha = 0.5) +
  geom_density(color = "darkblue", linewidth = 1) +
  labs(title = "Age Distribution", x = "Age", y = "Density") +
  scale_x_continuous(breaks = seq(0, max(cardio_df$age, na.rm = TRUE) + 10, by = 10)) +
  theme_bw()

@@

# Serum Cholesterol distribution

```{r}
p2 <- ggplot(subset(cardio_df, serumcholestrol > 0), aes(x = serumcholestrol)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "forestgreen", alpha = 0.5) +
  geom_density(color = "darkgreen", linewidth = 1) +
  labs(title = "Serum Cholesterol Distribution", 
       x = "Serum Cholesterol (mg/dL)", y = "Density") +
  theme_minimal()

@@

# Resting Blood Pressure distribution

```{r}
p3 <- ggplot(subset(cardio_df, restingBP > 0), aes(x = restingBP)) +
  geom_histogram(aes(y = ..density..), bins = 30, color = "black", fill = "maroon", alpha = 0.5) +
  geom_density(color = "darkred", linewidth = 1) +
  labs(title = "Resting Blood Pressure Distribution", 
       x = "Resting Blood Pressure (mmHG)", y = "Density") +
  theme_bw()

@@

# Boxplot for resting blood pressure by heart disease status

```{r}
p4 <- ggplot(cardio_df, aes(x = target, y = restingBP, fill = target)) +
  geom_boxplot() +
  labs(title = "Resting Blood Pressure by Heart Disease Status",
       x = "Heart Disease Status",
       y = "Resting Blood Pressure (mmHg)") +
  theme_minimal()

@@

# Violin plot for oldpeak by heart disease status

```{r}
p5 <- ggplot(cardio_df, aes(x = target, y = oldpeak, fill = target)) +
  geom_violin(trim = FALSE) +
  labs(title = "Oldpeak Distribution by Heart Disease Status",
       x = "Heart Disease Status",
       y = "Oldpeak") +
  theme_minimal()

@@

# Create the correlation matrix for selected numeric variables

```{r}
numeric_vars <- cardio_df %>% 
  select(age, restingBP, serumcholestrol, maxheartrate, oldpeak)
cor_matrix <- cor(numeric_vars, use = "pairwise.complete.obs")

### 2025-04-06 code/processing-code/processed file.qmd
Made a few changes to the readme files

# Load necessary libraries
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(here)           # For file paths
library(naniar)         # For gg_miss_var()
```

# PART 1: DATA IMPORT AND CLEANING
```{r}
# Import the dataset
cardio_df <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

@@

# PART 3: STATISTICAL ANALYSIS: DATA CLEANING, SUMMARIZATION, AND VISUALIZATION
```{r}
# Filter the data for individuals under 50 with normal cholesterol (<200 mg/dL) and create risk indicator variables for elevated blood pressure and ST depression.
cleaned_df <- cardio_df %>%
  filter(age < 50, serumcholestrol > 0, serumcholestrol < 200) %>%  
  mutate(
    # Create a binary variable for blood pressure:
    # "Elevated" if restingBP is between 120 and 129; otherwise, "Otherwise"
    elevated_BP = if_else(between(restingBP, 120, 129), "Elevated", "Otherwise"),

@@

# Save the processed data
```{r}
save_data_location <- here::here("data/processed-data/cleaned_df.rds")
saveRDS(cleaned_df, file = save_data_location)
```

>>>>

# Load necessary libraries

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(here)           # For file paths
library(naniar)         # For gg_miss_var()
```

# DATA IMPORT AND CLEANING

```{r}
# Import the dataset
cardio_df <- read_csv(here("data", "raw-data", "Cardiovascular_Disease_Dataset.csv"))

@@

# STATISTICAL ANALYSIS: DATA CLEANING, SUMMARIZATION, AND VISUALIZATION

```{r}
# Filter the data for individuals under 50 with normal cholesterol (<200 mg/dL) and create risk indicator variables for elevated blood pressure and ST depression.
cleaned_df <- cardio_df %>%
  filter(age < 50, serumcholestrol > 0, serumcholestrol < 200) %>%  
  mutate(
    # Create a binary variable for blood pressure:
    # "Elevated" if restingBP is between 120 and 129; otherwise, "Otherwise"
    elevated_BP = if_else(between(restingBP, 120, 129), "Elevated", "Otherwise"),

@@

# SAVE THE PROCESSED DATA

```{r}
save_data_location <- here::here("data/processed-data/cleaned_df.rds")
saveRDS(cleaned_df, file = save_data_location)
```

### 2025-04-06 products/manuscript/Manuscript.qmd
Made a few changes to the readme files

summary_df = skimr::skim(summary_stats)
print(summary_stats)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)

>>>>

summary_df = skimr::skim(summary_stats)
print(summary_stats)
# save to file
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary_df, file = summarytable_file)
